---
title: "Ch 1: Synthetic control [Step 3- model assessment, pvalues]"
author: "Molly Caldwell"
date: "01-January-2025"
output: html_document
---

```{r}
library(tidyverse)
```

# load in model output

Notes on treatment times:
For non-pronghorn species, treatment starts 2 hours before contact (time period 13) and contact is at time period 14.

For pronghorn, treatment starts 4 hours before contact (time period 12) and contact is at time period 14.

```{r}
# load in synthetic control models
syn_models <- readRDS("~/OneDrive - University of Wyoming/Documents/UWyo/PhD project/PhD_Chapter_1/Code output/gsynth_treated_output.rds")

# load in placebo models
plac_models <- readRDS("~/OneDrive - University of Wyoming/Documents/UWyo/PhD project/PhD_Chapter_1/Code output/gsynth_placebo_output.rds")
```

```{r}
# combine models into single dataframes
# first, remove models that did not run
syn_models <- map(syn_models, ~ discard(.x, ~ any(grepl("ERROR", unlist(.x)))))

plac_models <- map(plac_models, ~ discard(.x, ~ any(grepl("ERROR", unlist(.x)))))

# then, add a unique model ID for each model
syn_models <- map(syn_models, ~imap(.x, ~ mutate(.x[["result"]], 
                                                 model_id = paste(.x[["result"]]$species[[1]], 
                                                                  .x[["result"]]$species2[[1]],
                                                                  .x[["result"]]$season[[1]],
                                                                  .y, sep = "_"))))


plac_models <- map(plac_models, ~imap(.x, ~ mutate(.x[["result"]], 
                                                 model_id = paste(.x[["result"]]$species[[1]], 
                                                                  .x[["result"]]$species2[[1]],
                                                                  .x[["result"]]$season[[1]],
                                                                  .y, sep = "_"))))

# then, bind data rows 
syn_models <- bind_rows(syn_models)
plac_models <- bind_rows(plac_models)
```

```{r}
# add starting time period of treatment and species, species2, season ID for significance tests
syn_models <- syn_models %>% 
  group_by(model_id) %>%
  mutate(treat_start = min(time_period[post_treat == TRUE], na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(spp_seas = paste(species, species2, season, sep = "_"))

plac_models <- plac_models %>% 
  group_by(model_id) %>%
  mutate(treat_start = min(time_period[post_treat == TRUE], na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(spp_seas = paste(species, species2, season, sep = "_"))
```

```{r}
# filter out all models (placebo and treated) that have a treatment start time other than 13 (species other than pronghorn) or 12 (for pronghorn)
plac_models <- plac_models %>% 
  mutate(ts_flag = case_when(
    species != "pronghorn" & treat_start != 13 ~ 1,
    species == "pronghorn" & treat_start != 12 ~ 1,
    TRUE ~ 0
  )) %>% 
  filter(ts_flag == 0) %>% 
  select(-ts_flag)

syn_models <- syn_models %>% 
  mutate(ts_flag = case_when(
    species != "pronghorn" & treat_start != 13 ~ 1,
    species == "pronghorn" & treat_start != 12 ~ 1,
    TRUE ~ 0
  )) %>% 
  filter(ts_flag == 0) %>% 
  select(-ts_flag)
```

```{r}
#some pre-visualizing
d1 <- syn_models %>% filter(model_id == "pronghorn_elk_winter_20")
d2 <- plac_models %>% filter(spp_seas == "pronghorn_elk_winter")
  
ggplot() +
  geom_line(data = d2, aes(x = time_period, y = speed_eff, group = model_id),
            color = "grey", alpha = 0.2) +
  geom_line(data = d1, aes(x = time_period, y = speed_eff)) +
  geom_vline(data = d1, aes(xintercept = treat_start), linetype = 2) +
  theme_classic()

ggplot(d2 %>% filter(time_period == 16)) +
  geom_histogram(aes(x = speed_eff))
    
```

# Clean up models
## Treatment models

```{r}
# synthetic control - treated models
# remove models with MSE > 90% quantile of the pre-treatment fit
# first calculate mean square error of the pre-treatment fit
syn_models <- syn_models %>%
  mutate(speed_sq_error = speed_eff^2,
         angle_sq_error = angle_eff^2) %>% 
  na.omit(speed_sq_error, angle_sq_error) %>% 
  group_by(model_id, post_treat) %>% 
  mutate(speed_mse = sum(speed_sq_error) / nrow(.),
         angle_mse = sum(angle_sq_error) / nrow(.)) %>% 
  ungroup()
  
# visualize the MSE
hist(syn_models$speed_mse, breaks = 1000,
     xlim = c(0, 0.0001))

q_speed <- quantile(syn_models$speed_mse, 0.9)

hist(syn_models$angle_mse, breaks = 1000,
     xlim = c(0, 0.0001))

q_angle <- quantile(syn_models$angle_mse, 0.9)

# flag interactions with pre-treatment MSE > 90% quantiles
syn_models <- syn_models %>% 
  mutate(speed_mse_flag = if_else(speed_mse > q_speed & post_treat == FALSE, 1, 0),
         angle_mse_flag = if_else(angle_mse > q_angle & post_treat == FALSE, 1, 0)) %>% 
  group_by(model_id) %>%
  mutate(speed_mse_flag = if_else(sum(speed_mse_flag) > 0, 1, 0),
         angle_mse_flag = if_else(sum(angle_mse_flag) > 0, 1, 0)) %>% 
  ungroup()

syn_models <- syn_models %>% 
  mutate(speed_mse_flag = as.factor(speed_mse_flag),
         angle_mse_flag = as.factor(angle_mse_flag))

```

```{r}
# visualize flagged models
pair_seas <- unique(syn_models$spp_seas)

## speed
for(i in 1:length(pair_seas)){
  
p <- ggplot(syn_models %>% filter(spp_seas == pair_seas[[i]])) +
  geom_line(aes(x = time_period, y = speed_eff, group = model_id,
                color = speed_mse_flag, alpha = speed_mse_flag)) +
  geom_vline(aes(xintercept = treat_start), linetype = 2) +
  labs(title = paste(pair_seas[[i]], "speed effect", sep = " "))
  
print(p)
}

## angle
for(i in 1:length(pair_seas)){
  
p <- ggplot(syn_models %>% filter(spp_seas == pair_seas[[i]])) +
  geom_line(aes(x = time_period, y = angle_eff, group = model_id,
                color = angle_mse_flag, alpha = angle_mse_flag)) +
  geom_vline(aes(xintercept = treat_start), linetype = 2) +
  labs(title = paste(pair_seas[[i]], "angle effect", sep = " "))
  
print(p)
}

```

```{r}
# create separate speed and angle model dataframes with flagged MSE removed
speed_models <- syn_models %>% 
  filter(speed_mse_flag == 0) %>% 
  select(-starts_with("angle"))

angle_models <- syn_models %>% 
  filter(angle_mse_flag == 0) %>% 
  select(-starts_with("speed"))

# pull greatest MSE for speed and angle from filtered models
speed_MSE_cutoff <- max(speed_models$speed_mse)
angle_MSE_cutoff <- max(angle_models$angle_mse)
```

## Placebo models

```{r}
# synthetic control - placebo models
# remove models with MSE > treated models greatest MSE of the pre-treatment fit
# first calculate mean square error of the pre-treatment fit
plac_models <- plac_models %>%
  mutate(speed_sq_error = speed_eff^2,
         angle_sq_error = angle_eff^2) %>% 
  na.omit(speed_sq_error, angle_sq_error) %>% 
  group_by(model_id, post_treat) %>% 
  mutate(speed_mse = sum(speed_sq_error) / nrow(.),
         angle_mse = sum(angle_sq_error) / nrow(.)) %>% 
  ungroup()
  
# visualize the MSE
hist(plac_models$speed_mse, breaks = 1000,
     xlim = c(0, 0.0001))

hist(plac_models$angle_mse, breaks = 1000,
     xlim = c(0, 0.0001))

# flag interactions with pre-treatment MSE > MSE of filtered, treated models
plac_models <- plac_models %>% 
  mutate(speed_mse_flag = if_else(speed_mse > speed_MSE_cutoff & post_treat == FALSE, 1, 0),
         angle_mse_flag = if_else(angle_mse > angle_MSE_cutoff & post_treat == FALSE, 1, 0)) %>% 
  group_by(model_id) %>%
  mutate(speed_mse_flag = if_else(sum(speed_mse_flag) > 0, 1, 0),
         angle_mse_flag = if_else(sum(angle_mse_flag) > 0, 1, 0)) %>% 
  ungroup()

plac_models <- plac_models %>% 
  mutate(speed_mse_flag = as.factor(speed_mse_flag),
         angle_mse_flag = as.factor(angle_mse_flag))

```

```{r}
# visualize flagged models per species pair - season
pair_seas <- unique(plac_models$spp_seas)

## speed 
for(i in 1:length(pair_seas)){
  print(i)
p <- ggplot(plac_models %>% filter(spp_seas == pair_seas[[i]])) +
  geom_line(aes(x = time_period, y = speed_eff, group = model_id,
                color = speed_mse_flag, alpha = speed_mse_flag)) +
  geom_vline(aes(xintercept = treat_start), linetype = 2) +
  labs(title = paste(pair_seas[[i]], "speed effect", sep = " "))
  
print(p)
}

## angle
for(i in 1:length(pair_seas)){
  
p <- ggplot(plac_models %>% filter(spp_seas == pair_seas[[i]])) +
  geom_line(aes(x = time_period, y = angle_eff, group = model_id,
                color = angle_mse_flag, alpha = angle_mse_flag)) +
  geom_vline(aes(xintercept = treat_start), linetype = 2) +
  labs(title = paste(pair_seas[[i]], "angle effect", sep = " "))
  
print(p)
}  

```

```{r}
# create separate speed and angle model dataframes with flagged MSE removed
speed_plac <- plac_models %>% 
  filter(speed_mse_flag == 0) %>% 
  select(-starts_with("angle"))

angle_plac <- plac_models %>% 
  filter(angle_mse_flag == 0) %>% 
  select(-starts_with("speed"))
```

# Assess control vs. placebo significance

```{r}
# create datasets of combined control and placebo models per speed and angle
speed_models$mod_type = "treated"
speed_plac$mod_type = "placebo"

speed_comb <- bind_rows(speed_models, speed_plac)

angle_models$mod_type = "treated"
angle_plac$mod_type = "placebo"

angle_comb <- bind_rows(angle_models, angle_plac)
```

```{r}
# visualize placebo post-treatment effects vs. treated effects (per species & season)
spp_id <- unique(speed_comb$spp_seas)

for(i in 1:length(spp_id)){
  p_speed <- speed_comb %>% 
              filter(spp_seas == spp_id[[i]] & time_period == treat_start) %>%
    group_by(mod_type) %>% 
    mutate(avg_speed_eff = mean(speed_eff)) %>% 
    ungroup() %>% 
    ggplot() +
    geom_histogram(aes(x = speed_eff, y = after_stat(density), fill = mod_type,
                       color = mod_type), 
                   alpha = 0.7, bins = 60) +
    geom_vline(aes(xintercept = avg_speed_eff, color = mod_type), 
               linewidth = 1.5, linetype = 2) +
    labs(title = paste0("Speed treatment effect vs. placebo: ", spp_id[[i]])) +
    theme_classic()
  
   p_angle <- angle_comb %>% 
              filter(spp_seas == spp_id[[i]] & time_period == treat_start) %>%
    group_by(mod_type) %>% 
    mutate(avg_angle_eff = mean(angle_eff)) %>% 
    ungroup() %>%  
    ggplot() +
    geom_histogram(aes(x = angle_eff, y = after_stat(density), fill = mod_type,
                       color = mod_type), 
                   alpha = 0.7, bins = 60) +
    geom_vline(aes(xintercept = avg_angle_eff, color = mod_type), 
               linewidth = 1.5, linetype = 2) +
    labs(title = paste0("Angle treatment effect vs. placebo: ", spp_id[[i]])) +
    theme_classic()
   
   print(p_speed)
   print(p_angle)
}

```

## Calculate p-values

```{r}
# function to calculate p-value based on treated time
calc_pvalue_treat <- function(treat_data, plac_data, time_sel){
  
  # filter placebo and treated data to time period of interest
  plac_data <- plac_data %>% filter(time_period == time_sel)
  treat_data <- treat_data %>%  filter(time_period == time_sel)
  
  # add eff column dependent on whether its speed or angle
  if(any(grepl("speed_eff", names(treat_data)))){
    treat_data$eff <- treat_data$speed_eff
    plac_data$eff <- plac_data$speed_eff
  }else{
    treat_data$eff <- treat_data$angle_eff
    plac_data$eff <- plac_data$angle_eff
  }
  
  # then, calculate the percent rank and pvalue of each treated effect
   dat <- treat_data %>% 
     rowwise() %>% 
    mutate(eff_perc_rank = sum(plac_data$eff < eff) / nrow(plac_data)) %>% 
    ungroup() %>% 
    mutate(eff_pvalue = if_else(eff_perc_rank >= 0.5, 1 - eff_perc_rank,
                                eff_perc_rank)) %>% 
     select(-eff, -eff_perc_rank)
   
   # rename pvalues dependent on whether its speed or angle
  if(any(grepl("speed_eff", names(treat_data)))){
    dat <- dat %>% 
      rename(speed_pvalue = eff_pvalue)
    
  }else{
    dat <- dat %>% 
      rename(angle_pvalue = eff_pvalue)
  }
  
  return(dat)
}

```

```{r}
# first, split treated and placebo datasets into spp-seas
speed_models <- split(speed_models, speed_models$spp_seas)
angle_models <- split(angle_models, angle_models$spp_seas)

speed_plac <- split(speed_plac, speed_plac$spp_seas)
angle_plac <- split(angle_plac, angle_plac$spp_seas)

# time periods of interest for pvalues:
# time of treatment (time before contact)
# time of contact
# time after contact
# 2 time periods after contact
# 13, 14, 15, 16 for non-pronghorn, and 12, 14, 16, 18 for pronghorn
times <- c(12:18)
# iterate the p-value function across each spp-seas and loop by time period of interest
speed_pval <- list()
angle_pval <- list()

for(i in 1:length(times)){
  print(i)
  
  speed_pval[[i]] <- map2_dfr(speed_models, speed_plac, 
                     ~calc_pvalue_treat(.x, .y, time_sel = times[[i]]))
  
  angle_pval[[i]] <- map2_dfr(angle_models, angle_plac, 
                     ~calc_pvalue_treat(.x, .y, time_sel = times[[i]]))
}

# combine lists into one dataframe
speed_pval <- bind_rows(speed_pval)

angle_pval <- bind_rows(angle_pval)

# join p-value data to full model data
speed_models <- bind_rows(speed_models)
speed_models <- left_join(speed_models, speed_pval)
angle_models <- bind_rows(angle_models)
angle_models <- left_join(angle_models, angle_pval)
```


```{r}
# summarize significant models
# first, combine speed and angle models
treat_models <- full_join(speed_models, angle_models)

# add flag for significant models at each time of interest
treat_models <- treat_models %>% 
  mutate(
    before_contact_sig = case_when(
           species == "pronghorn" & time_period == 12 & speed_pvalue <= 0.05 & 
             angle_pvalue <= 0.05 ~ "both",
           species != "pronghorn" & time_period == 13 & speed_pvalue <= 0.05 &
             angle_pvalue <= 0.05 ~ "both",
           species == "pronghorn" & time_period == 12 & speed_pvalue <= 0.05 ~ "speed",
           species != "pronghorn" & time_period == 13 & speed_pvalue <= 0.05 ~ "speed",
           species == "pronghorn" & time_period == 12 & angle_pvalue <= 0.05 ~ "angle",
           species != "pronghorn" & time_period == 13 & angle_pvalue <= 0.05 ~ "angle",
           TRUE ~ "none"), 
    contact_sig = case_when(
          time_period == 14 & speed_pvalue <= 0.05 & angle_pvalue <= 0.05 ~ "both",
           time_period == 14 & speed_pvalue <= 0.05 ~ "speed",
           time_period == 14 & angle_pvalue <= 0.05 ~ "angle",
           TRUE ~ "none"),
    after_contact_1_sig = case_when(
      species == "pronghorn" & time_period == 16 & speed_pvalue <= 0.05 & 
             angle_pvalue <= 0.05 ~ "both",
           species != "pronghorn" & time_period == 15 & speed_pvalue <= 0.05 &
             angle_pvalue <= 0.05 ~ "both",
           species == "pronghorn" & time_period == 16 & speed_pvalue <= 0.05 ~ "speed",
           species != "pronghorn" & time_period == 15 & speed_pvalue <= 0.05 ~ "speed",
           species == "pronghorn" & time_period == 16 & angle_pvalue <= 0.05 ~ "angle",
           species != "pronghorn" & time_period == 15 & angle_pvalue <= 0.05 ~ "angle",
           TRUE ~ "none"),
    after_contact_2_sig = case_when(
           species == "pronghorn" & time_period == 18 & speed_pvalue <= 0.05 & 
             angle_pvalue <= 0.05 ~ "both",
           species != "pronghorn" & time_period == 16 & speed_pvalue <= 0.05 &
             angle_pvalue <= 0.05 ~ "both",
           species == "pronghorn" & time_period == 18 & speed_pvalue <= 0.05 ~ "speed",
           species != "pronghorn" & time_period == 16 & speed_pvalue <= 0.05 ~ "speed",
           species == "pronghorn" & time_period == 18 & angle_pvalue <= 0.05 ~ "angle",
           species != "pronghorn" & time_period == 16 & angle_pvalue <= 0.05 ~ "angle",
           TRUE ~ "none")) %>% 
  group_by(model_id) %>% 
  mutate(before_contact_sig = case_when(
    any(before_contact_sig == "speed") ~ "speed",
    any(before_contact_sig == "angle") ~ "angle",
    any(before_contact_sig == "both") ~ "both",
    TRUE ~ "none"),
    contact_sig = case_when(
    any(contact_sig == "speed") ~ "speed",
    any(contact_sig == "angle") ~ "angle",
    any(contact_sig == "both") ~ "both",
    TRUE ~ "none"),
    after_contact_1_sig = case_when(
    any(after_contact_1_sig == "speed") ~ "speed",
    any(after_contact_1_sig == "angle") ~ "angle",
    any(after_contact_1_sig == "both") ~ "both",
    TRUE ~ "none"),
    after_contact_2_sig = case_when(
    any(after_contact_2_sig == "speed") ~ "speed",
    any(after_contact_2_sig == "angle") ~ "angle",
    any(after_contact_2_sig == "both") ~ "both",
    TRUE ~ "none")) %>% 
  ungroup() 
```

```{r}
# remove duplicates
treat_models <- treat_models %>% 
  select(-model_id)

treat_models <- treat_models %>% 
  mutate(across(where(is.numeric), ~round(.,5)))

treat_models <- treat_models[!duplicated(treat_models),]

# save treat data with p vals
saveRDS(treat_models, "~/OneDrive - University of Wyoming/Documents/UWyo/PhD project/PhD_Chapter_1/Code output/syn_models_output_pval.rds")
```


```{r}
sig_prop <- treat_models %>%
  distinct(id1, id2, date_int, .keep_all = TRUE) %>% 
  group_by(spp_seas, species, after_contact_1_sig) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(spp_seas, species) %>%
  mutate(proportion = count / sum(count))

# plot number of significant models per species & season
ggplot(sig_prop %>% filter(after_contact_1_sig != "none"), 
       aes(x = spp_seas, y = proportion, fill = after_contact_1_sig)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~species, scales = "free") +
  coord_flip() +
  theme_classic() 
```

## Assess p-values with Bayesian bootstrap of placebos

```{r}
# function for bayesian bootstrap resample
bayesian_bootstrap <- function(plac_data, time_sel, n_resamples = 1000) {
  
  # filter to time period of interest
  plac_data <- plac_data %>% 
    filter(time_period == time_sel)
  
  # Length of the data
  n <- nrow(plac_data)
  
  # put condition to continue only if nrow plac_data > 0 (to skip models without certain time periods for pronghorn)
  
  if(n > 0){
  # Generate Dirichlet weights
  weights <- gtools::rdirichlet(1, rep(1, n))
  
  # create dataframe of resampled effects
  if(any(grepl("speed_eff", names(plac_data)))){
    # resample data with the weights
  speed_eff <- sample(plac_data$speed_eff, 
                      size = n_resamples, 
                      replace = TRUE, 
                      prob = weights)
  
  dat <- data.frame(spp_seas = plac_data$spp_seas[[1]], 
                    time_period = time_sel,
                    speed_eff = speed_eff)
  }else{
    # resample data with the weights
  angle_eff <- sample(plac_data$angle_eff, 
                      size = n_resamples, 
                      replace = TRUE, 
                      prob = weights)
  
  dat <- data.frame(spp_seas = plac_data$spp_seas[[1]], 
                    time_period = time_sel,
                    angle_eff = angle_eff)
  }
  return(dat)
  }
}

```

```{r}
# time periods of interest for pvalues:
# time of treatment (time before contact)
# time of contact
# time after contact
# 2 time periods after contact
# 13, 14, 15, 16 for non-pronghorn, and 12, 14, 16, 18 for pronghorn
times <- c(12:18)

# iterate the bayesian bootstrap across placebo models and time period of interest
speed_boot <- list()
angle_boot <- list()

for(i in 1:length(times)){
  print(i)
  
  speed_boot[[i]] <- map_dfr(speed_plac, 
                        ~ bayesian_bootstrap(.x, time_sel = times[[i]]))
  
  angle_boot[[i]] <- map_dfr(angle_plac, 
                        ~ bayesian_bootstrap(.x, time_sel = times[[i]]))
}

speed_boot <- bind_rows(speed_boot)
angle_boot <- bind_rows(angle_boot)

# split boot data into list by spp_seas
speed_boot <- split(speed_boot, speed_boot$spp_seas)
angle_boot <- split(angle_boot, angle_boot$spp_seas)
```

```{r}
# remove pvalues if already in model data
speed_models <- speed_models %>% 
  select(-speed_pvalue)

angle_models <- angle_models %>% 
  select(-angle_pvalue)

# split model data into list by spp_seas
speed_models <- split(speed_models, speed_models$spp_seas)
angle_models <- split(angle_models, angle_models$spp_seas)

# iterate the p-value function across each spp-seas and loop by time period of interest
speed_pval_boot <- list()
angle_pval_boot <- list()

for(i in 1:length(times)){
  print(i)
  
  speed_pval_boot[[i]] <- map2_dfr(speed_models, speed_boot, 
                     ~calc_pvalue_treat(.x, .y, time_sel = times[[i]]))
  
  angle_pval_boot[[i]] <- map2_dfr(angle_models, angle_boot, 
                     ~calc_pvalue_treat(.x, .y, time_sel = times[[i]]))
}

# combine lists into one dataframe
speed_pval_boot <- bind_rows(speed_pval_boot)

angle_pval_boot <- bind_rows(angle_pval_boot)

# join p-value data to full model data
speed_models <- bind_rows(speed_models)
speed_models <- left_join(speed_models, speed_pval_boot)
angle_models <- bind_rows(angle_models)
angle_models <- left_join(angle_models, angle_pval_boot)
```

```{r}
# save bootstrapped placebo data
speed_boot_df <- bind_rows(speed_boot)
angle_boot_df <- bind_rows(angle_boot)

boot_df <- speed_boot_df
boot_df$angle_eff <- speed_boot_df$speed_eff

saveRDS(boot_df, "~/OneDrive - University of Wyoming/Documents/UWyo/PhD project/PhD_Chapter_1/Code output/placebo_bayesboot_effects.rds")
```

```{r}
# summarize significant models
# first, combine speed and angle models
treat_models_boot <- full_join(speed_models, angle_models)

# add flag for significant models at each time of interest
treat_models_boot <- treat_models_boot %>% 
  mutate(
    before_contact_sig = case_when(
           species == "pronghorn" & time_period == 12 & speed_pvalue <= 0.05 & 
             angle_pvalue <= 0.05 ~ "both",
           species != "pronghorn" & time_period == 13 & speed_pvalue <= 0.05 &
             angle_pvalue <= 0.05 ~ "both",
           species == "pronghorn" & time_period == 12 & speed_pvalue <= 0.05 ~ "speed",
           species != "pronghorn" & time_period == 13 & speed_pvalue <= 0.05 ~ "speed",
           species == "pronghorn" & time_period == 12 & angle_pvalue <= 0.05 ~ "angle",
           species != "pronghorn" & time_period == 13 & angle_pvalue <= 0.05 ~ "angle",
           TRUE ~ "none"), 
    contact_sig = case_when(
          time_period == 14 & speed_pvalue <= 0.05 & angle_pvalue <= 0.05 ~ "both",
           time_period == 14 & speed_pvalue <= 0.05 ~ "speed",
           time_period == 14 & angle_pvalue <= 0.05 ~ "angle",
           TRUE ~ "none"),
    after_contact_1_sig = case_when(
      species == "pronghorn" & time_period == 16 & speed_pvalue <= 0.05 & 
             angle_pvalue <= 0.05 ~ "both",
           species != "pronghorn" & time_period == 15 & speed_pvalue <= 0.05 &
             angle_pvalue <= 0.05 ~ "both",
           species == "pronghorn" & time_period == 16 & speed_pvalue <= 0.05 ~ "speed",
           species != "pronghorn" & time_period == 15 & speed_pvalue <= 0.05 ~ "speed",
           species == "pronghorn" & time_period == 16 & angle_pvalue <= 0.05 ~ "angle",
           species != "pronghorn" & time_period == 15 & angle_pvalue <= 0.05 ~ "angle",
           TRUE ~ "none"),
    after_contact_2_sig = case_when(
           species == "pronghorn" & time_period == 18 & speed_pvalue <= 0.05 & 
             angle_pvalue <= 0.05 ~ "both",
           species != "pronghorn" & time_period == 16 & speed_pvalue <= 0.05 &
             angle_pvalue <= 0.05 ~ "both",
           species == "pronghorn" & time_period == 18 & speed_pvalue <= 0.05 ~ "speed",
           species != "pronghorn" & time_period == 16 & speed_pvalue <= 0.05 ~ "speed",
           species == "pronghorn" & time_period == 18 & angle_pvalue <= 0.05 ~ "angle",
           species != "pronghorn" & time_period == 16 & angle_pvalue <= 0.05 ~ "angle",
           TRUE ~ "none")) %>% 
  group_by(model_id) %>% 
  mutate(before_contact_sig = case_when(
    any(before_contact_sig == "speed") ~ "speed",
    any(before_contact_sig == "angle") ~ "angle",
    any(before_contact_sig == "both") ~ "both",
    TRUE ~ "none"),
    contact_sig = case_when(
    any(contact_sig == "speed") ~ "speed",
    any(contact_sig == "angle") ~ "angle",
    any(contact_sig == "both") ~ "both",
    TRUE ~ "none"),
    after_contact_1_sig = case_when(
    any(after_contact_1_sig == "speed") ~ "speed",
    any(after_contact_1_sig == "angle") ~ "angle",
    any(after_contact_1_sig == "both") ~ "both",
    TRUE ~ "none"),
    after_contact_2_sig = case_when(
    any(after_contact_2_sig == "speed") ~ "speed",
    any(after_contact_2_sig == "angle") ~ "angle",
    any(after_contact_2_sig == "both") ~ "both",
    TRUE ~ "none")) %>% 
  ungroup() 
```

```{r}
# remove duplicates
treat_models_boot<- treat_models_boot %>% 
  select(-model_id)

treat_models_boot <- treat_models_boot %>% 
  mutate(across(where(is.numeric), ~round(.,5)))

treat_models_boot <- treat_models_boot[!duplicated(treat_models_boot),]

# save treat data with p vals
saveRDS(treat_models_boot, "~/OneDrive - University of Wyoming/Documents/UWyo/PhD project/PhD_Chapter_1/Code output/syn_models_output_pval_bayesboot.rds")
```

```{r}
sig_prop <- treat_models_boot %>%
  distinct(id1, id2, date_int, .keep_all = TRUE) %>% 
  group_by(spp_seas, species, contact_sig) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(spp_seas, species) %>%
  mutate(proportion = count / sum(count))

# plot number of significant models per species & season
ggplot(sig_prop %>% filter(contact_sig != "none"), 
       aes(x = spp_seas, y = proportion, fill = contact_sig)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~species, scales = "free_y") +
  coord_flip() +
  theme_classic() +
  labs(y = "Proportion significant effects", x = "")
```

## compare boot pvalues to placebo data pvalues

```{r}
sig_plac <- treat_models %>%
  distinct(model_id, .keep_all = TRUE) %>% 
  group_by(spp_seas, species, after_contact_1_sig) %>%
  summarise(count = n(), .groups = "drop") %>% 
  group_by(spp_seas, species) %>%
  mutate(proportion = count / sum(count)) %>% 
  ungroup() %>% 
  mutate(pval_type = "placebo")

sig_boot <- treat_models_boot %>%
  distinct(model_id, .keep_all = TRUE) %>% 
  group_by(spp_seas, species, after_contact_1_sig) %>%
  summarise(count = n(), .groups = "drop") %>% 
  group_by(spp_seas, species) %>%
  mutate(proportion = count / sum(count)) %>% 
  ungroup() %>% 
  mutate(pval_type = "bootstrap")

sig_comp <- bind_rows(sig_plac, sig_boot) 

ggplot(sig_comp %>% filter(after_contact_1_sig != "none"), 
       aes(x = spp_seas, y = proportion, fill = pval_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~after_contact_1_sig, scales = "free") +
  coord_flip() +
  theme_classic() 
```

```{r}
# bind together bootstrap values
angle_boot <- bind_rows(angle_boot)
speed_boot <- bind_rows(speed_boot)

# visualize bootstrapped placebo post-treatment effects vs. treated effects (per species & season)
spp_id <- unique(speed_models$spp_seas)

for(i in 1:length(spp_id)){
  
  print(i)
  
  if(grepl("^pronghorn", spp_id[[i]])){
    time = 16
  }else{
    time = 15
  }
  
  speed <- speed_models %>% 
    filter(spp_seas == spp_id[[i]] & time_period == time) %>% 
    mutate(avg_speed_eff = mean(speed_eff))
  
   angle <- angle_models %>% 
    filter(spp_seas == spp_id[[i]] & time_period == time) %>% 
    mutate(avg_angle_eff = mean(angle_eff))
   
   speed_boot_f <- speed_boot  %>% 
    filter(spp_seas == spp_id[[i]] & time_period == time) %>% 
    mutate(avg_speed_eff = mean(speed_eff))
   
   angle_boot_f <- angle_boot %>% 
    filter(spp_seas == spp_id[[i]] & time_period == time) %>% 
    mutate(avg_angle_eff = mean(angle_eff))
  
  p_speed <- 
    ggplot() +
    geom_histogram(data = speed_boot_f,
      aes(x = speed_eff, y = after_stat(density)), 
                   fill = "darkgrey", alpha = 0.7, bins = 60) +
    geom_vline(data = speed_boot_f,
               aes(xintercept = avg_speed_eff), 
               color = "black", linewidth = 1, linetype = 2) +
    geom_histogram(data = speed,
      aes(x = speed_eff, y = after_stat(density)), 
                   fill = "blue", alpha = 0.5, bins = 60) +
    geom_vline(data = speed,
               aes(xintercept = avg_speed_eff), 
               color = "blue", linewidth = 1, linetype = 2) +
    labs(title = paste0("Speed treatment effect vs. placebo: ", spp_id[[i]])) +
    theme_classic()
  
   p_angle <- ggplot() +
    geom_histogram(data = angle_boot_f,
      aes(x = angle_eff, y = after_stat(density)), 
                   fill = "darkgrey", alpha = 0.7, bins = 60) +
    geom_vline(data = angle_boot_f,
               aes(xintercept = avg_angle_eff), 
               color = "black", linewidth = 1, linetype = 2) +
    geom_histogram(data = angle,
      aes(x = angle_eff, y = after_stat(density)), 
                   fill = "blue", alpha = 0.5, bins = 60) +
    geom_vline(data = angle,
               aes(xintercept = avg_angle_eff), 
               color = "blue", linewidth = 1, linetype = 2) +
    labs(title = paste0("Angle treatment effect vs. placebo: ", spp_id[[i]])) +
    theme_classic()
   
   print(p_speed)
   print(p_angle)
}

```







